---
title: "Kenya Census"
format:
  html:
    theme: cosmo
    page-layout: full
    toc: false
    css: styles.css
editor: visual
---

```{r load, echo = F, message = F, warning=F}
pacman::p_load(dplyr, here, tidyverse, readr, readxl, janitor, gtsummary, lubridate, gtsummary)
source(here("code", "1_setup.R"))

```

# Where respondent seek treatment first

```{r seektreat, echo = F}
health_vars <- c("health_seek_first_clean")

health_single <- make_tbl_by_comm(
  clean,
  health_vars,
  label_list = list(
    health_seek_first_clean ~ "Where do you seek treatment first?"
  )
)

health_single   %>% 
  as_gt() %>%
  opt_row_striping()
```

## Other treatment options

```{r seektreat_other, echo = F}
clean %>%
  filter(health_seek_first_clean == "Other",
         !is.na(health_seek_first_other),
         health_seek_first_other != "") %>%
  count(geo_community, health_seek_first_other, name = "Count") %>%
  arrange(desc(Count)) %>%   # <-- sort here
  gt::gt() %>%
  gt::cols_label(
    geo_community = "Community",
    health_seek_first_other = "Other responses"
  ) %>%
  gt::fmt_number(columns = Count, decimals = 0) %>%
  gt::opt_row_striping()
```

# Health questions

```{r healthvars, echo = F}
health_vars <- c("health_travel_time","health_chp_visit_monthly","health_chp_econ_beneficiary", "health_sick_3mo","health_sick_treat_seek","health_program_aware"
)

health_single <- make_tbl_by_comm(clean, health_vars)
health_single   %>% 
  as_gt() %>%
  opt_row_striping()
```

## Other health programs

**Note on response standardization**

The following health program responses were consolidated to improve readability while preserving meaning:

-   Responses mentioning **410 Bridge or 410-supported activities** (e.g., *410 Bridge*, *410 health program(s)*, *410 medical camp*, *410 health camping*, etc.) were grouped as **"410-related."**

-   Variations of **medical camps** (e.g., *medical camp*, *medical camps*, *medical camp's*) were grouped as **"Medical camp."**

-   Variations of **free medical camps** (e.g., *free medical camp*, *free medical camps*, *free medical*) were grouped as **"Free medical camp."**

-   Responses mentioning **deworming activities** (e.g., *deworming*, *deworming of pupils*, *pupils deworming*, *deworming by county government*, etc.) were grouped as **"Deworming."**

-   Variations of **vaccination activities** (e.g., *vaccination*, *vaccinations*, *vaccines*, *child vaccination*) were grouped as **"Vaccination."**

-   Variations of **Community Health Promoter (CHP) activities** (e.g., *CHP*, *Chp*, *CHPs*) were grouped as **"CHP."**

-   References to **CHP trainings or refresher activities** (e.g., *CHPs training*, *refresher course*, *training of CHPs*, *CHPs train on polio management*, etc.) were grouped as **"CHP training."**

-   Variations referring to **distribution of mosquito nets** (e.g., *net distribution*, *nets distribution*, *net supply*, *giving out nets*, *mosquito nets*) were grouped as **"Net distribution."**

-   Variations referring to **health screening activities** (e.g., *cancer screening*, *cervical cancer screening*, *screening*) were grouped as **"Screening."**

-   Variations referring to **community health outreach or organizations** (e.g., *community outreach*, *community-based health organization*, county outreach programs) were grouped as **"Community outreach."**

-   Variations of **Dignity for Girls programming** (e.g., *dignity for girls*, *dignity for girls and boys*) were grouped as **"Dignity for girls."**

-   Variations referring to **health insurance registration** (e.g., *SHA registration*, *Social Health Insurance Fund registration*) were grouped as **"Health insurance registration."**

-   Variations referring to **maternal health programs** (e.g., *maternal health*, training for expectant mothers) were grouped as **"Maternal health."**

-   Entries indicating **no program mentioned** (e.g., *No*, *None*) were grouped as **"None."**

-   All other responses were retained as originally recorded.

```{r other_health_programs, echo = F}
health_prog_merged <- clean %>%
  filter(!is.na(health_program_list), health_program_list != "") %>%
  separate_rows(health_program_list, sep = "[;,]") %>%
  mutate(program_raw = str_squish(health_program_list)) %>%
  filter(program_raw != "") %>%
  mutate(
    program_group = case_when(
       str_detect(program_raw, regex("^aka\\s*sivitali\\.?ulu$", ignore_case = TRUE)) ~ "Aka sivitali ulu",

    str_detect(program_raw, regex("^children\\s+vacation$", ignore_case = TRUE)) ~ "Vaccination",

    str_detect(program_raw, regex("^chps?\\s*imitiatives$|^chps?\\s*initiatives$", ignore_case = TRUE)) ~ "CHP",
    str_detect(program_raw, regex("^community\\s+health\\s+promoters?$", ignore_case = TRUE)) ~ "CHP",

    
      # 410-related (any mention of 410 or 410 Bridge)
      str_detect(program_raw, regex("\\b410\\b|410\\s*bridge|^410", ignore_case = TRUE)) ~ "410-related",

      # Keep FREE medical camp separate from Medical camp
      str_detect(program_raw, regex("free\\s+medical\\s+camp", ignore_case = TRUE)) ~ "Free medical camp",
      str_detect(program_raw, regex("medical\\s+camp", ignore_case = TRUE)) ~ "Medical camp",

      # Deworming (anything that mentions deworming)
      str_detect(program_raw, regex("deworm", ignore_case = TRUE)) ~ "Deworming",

      # Vaccination/vaccinations/vaccines
      str_detect(program_raw, regex("^vaccin", ignore_case = TRUE)) ~ "Vaccination",

      # CHP variants (CHP, Chps, chp, etc.)
      str_detect(program_raw, regex("^chp(s)?$|\\bchp\\b", ignore_case = TRUE)) ~ "CHP",
      
      # Nets
  str_detect(program_raw, regex("net", TRUE)) ~ "Net distribution",

  # Screening (keep health screening separate if you want)
  str_detect(program_raw, regex("cancer screening|cervical cancer screening|\\bscreening\\b", TRUE)) ~ "Screening",

  # CHP training bucket (optional but helpful)
  str_detect(program_raw, regex("refresher|train|training|first aid|polio management|doctors association", TRUE)) &
    str_detect(program_raw, regex("\\bchp", TRUE)) ~ "CHP training",

  # Outreach / community org
  str_detect(program_raw, regex("outreach|community based health organisation", TRUE)) ~ "Community outreach",

  # Dignity (optional)
  str_detect(program_raw, regex("dignity for girls", TRUE)) ~ "Dignity for girls",

  # Insurance (optional)
  str_detect(program_raw, regex("sha registration|social health insurance", TRUE)) ~ "Health insurance registration",

  # Maternal (optional)
  str_detect(program_raw, regex("maternal|expectant mothers", TRUE)) ~ "Maternal health",

  # None / No (optional)
  str_detect(program_raw, regex("^no$|^none$", TRUE)) ~ "None",

      TRUE ~ str_to_sentence(program_raw)  # keep others readable but unmerged
    )
  ) %>%
  count(program_group, name = "Count", sort = TRUE)

health_prog_merged %>%
  gt() %>%
  tab_header(title = "Health programs mentioned") %>%
  cols_label(program_group = "Program", Count = "Count") %>%
  fmt_number(columns = Count, decimals = 0) %>%
  opt_row_striping()


```

## Other health programs (disaggregated by community)

```{r other_health_comm, echo = F}
health_prog_by_comm <- clean %>%
  filter(!is.na(health_program_list), health_program_list != "") %>%
  separate_rows(health_program_list, sep = "[;,]") %>%
  mutate(program_raw = str_squish(health_program_list)) %>%
  filter(program_raw != "") %>%
  mutate(
    program_group = case_when(
       str_detect(program_raw, regex("^aka\\s*sivitali\\.?ulu$", ignore_case = TRUE)) ~ "Aka sivitali ulu",

    str_detect(program_raw, regex("^children\\s+vacation$", ignore_case = TRUE)) ~ "Vaccination",

    str_detect(program_raw, regex("^chps?\\s*imitiatives$|^chps?\\s*initiatives$", ignore_case = TRUE)) ~ "CHP",
    str_detect(program_raw, regex("^community\\s+health\\s+promoters?$", ignore_case = TRUE)) ~ "CHP",
      # 410-related
      str_detect(program_raw, regex("\\b410\\b|410\\s*bridge|^410", ignore_case = TRUE)) ~ "410-related",

      # keep FREE medical camp separate from Medical camp
      str_detect(program_raw, regex("free\\s+medical\\s+camp", ignore_case = TRUE)) ~ "Free medical camp",
      str_detect(program_raw, regex("medical\\s+camp", ignore_case = TRUE)) ~ "Medical camp",

      # deworming
      str_detect(program_raw, regex("deworm", ignore_case = TRUE)) ~ "Deworming",

      # vaccination
      str_detect(program_raw, regex("^vaccin", ignore_case = TRUE)) ~ "Vaccination",

      # CHP training (must come before CHP)
      str_detect(program_raw, regex("refresher|train|training|first aid|polio management|doctors association", TRUE)) &
        str_detect(program_raw, regex("\\bchp", TRUE)) ~ "CHP training",

      # CHP variants
      str_detect(program_raw, regex("^chp(s)?$|\\bchp\\b", ignore_case = TRUE)) ~ "CHP",

      # nets
      str_detect(program_raw, regex("net", ignore_case = TRUE)) ~ "Net distribution",

      # screening
      str_detect(program_raw, regex("cancer screening|cervical cancer screening|\\bscreening\\b", ignore_case = TRUE)) ~ "Screening",

      # outreach
      str_detect(program_raw, regex("outreach|community based health organisation", ignore_case = TRUE)) ~ "Community outreach",

      # dignity
      str_detect(program_raw, regex("dignity for girls", ignore_case = TRUE)) ~ "Dignity for girls",

      # insurance
      str_detect(program_raw, regex("sha registration|social health insurance", ignore_case = TRUE)) ~ "Health insurance registration",

      # maternal
      str_detect(program_raw, regex("maternal|expectant mothers", ignore_case = TRUE)) ~ "Maternal health",

      # none/no
      str_detect(program_raw, regex("^no$|^none$", ignore_case = TRUE)) ~ "None",

      TRUE ~ str_to_sentence(program_raw)
    )
  ) %>%
  count(program_group, geo_community, name = "Count") %>%
  pivot_wider(
    names_from = geo_community,
    values_from = Count,
    values_fill = 0
  ) %>%
  arrange(desc(rowSums(select(., where(is.numeric)))))

health_prog_by_comm %>%
  slice_head(n = 30) %>%
  gt() %>%
  tab_header(title = "Health programs mentioned — by community") %>%
  cols_label(program_group = "Program") %>%
  fmt_number(columns = where(is.numeric), decimals = 0) %>%
  opt_row_striping()
```

## 
